### 2019-03-10 papers/journal/encoding.tex
Improvement in Univ condition section 6.3

\textbf{Constraints for stability condition}
%
We use \texttt{NOracle} to represent the required path to encode the stability condition in VTS.
%
A naive implementation is shown in~\ref{eq:reachbmc1} where \texttt{NOracle}$(m, i, j, w)$ represent the presence of a $m$-$path$ from node $i$ to node $j$ of maximum $w$ length.
%there is a edge between $i^{\prime \prime}$ and01
%$(i \neq i^{\prime \prime})$ with $m$ present, and j is $m$-reachable from $i'$ in less than $p-1$ steps.
%
% \begin{align}
%   NOracle (m, i, j, w) = \exists \, path[h1, ..., hn]: \\ 2 \leq |path| \leq  w + 2  \, \land h1 = i \, \land  hn = j \, \land \\ 
%   \bigvee\limits_{k = 1}^{|path|- 1} \bigvee\limits_{q}  E[hk] [hk+1][q][m] 
%   \tag{S1}\label{eq:reach1}
% \end{align}
%The NOracle can be thought of as a disjunction of $m$-path of increment length. 
%
It performs enumeration of every $m$-$path$, $m$-${paths}_i^j$, of at most $w$ length. 
%
Starting with exploring $m$-${paths}_i^j$ of length 1, a direct edge between $i$ and $j$ with molecule $m$ present on it, until every $m$-${paths}_i^j$ of length $w-1$ is explored.
%by performing series of node hops till final node destination j is reached.
%Starting with $m$-${paths}_i^j$ of length 1 which is a direct edge between $i$ and $j$ with molecule $m$ present on the edge or performing series of node hops till final node destination j is reached.

@@

\subsection{Universal condition}
%
Recall that the specification for universal condition~\ref{eq:univcond} requires quantifier alternation. 
%
Most SAT solvers can not efficiently handle the quantifier in the formulas. 
% 
Although, it is possible to encode the universal condition~\ref{eq:univcond} as a SAT or SMT problem.
%
There is always a danger of a blow-up in the formula size, making it harder for the SAT solver to handle. 
%Which may blow up the formula size for the solver to solve.
% 
As an experiment we have implemented a trivial SMT version of the universal problem extending the current SMT encodings.
%
The basic idea is to conjunct the formula for each combination value of the quantified variable. 
%

\input{satqbf-table}
%
The table~\ref{tab:satqbf-graph}, shows results of search result of universal condition for N = 4, 3-connected graphs. The SAT version was not able to scale it to the desired results and either timed-out (T/O) or ran out of memory (M/O).
%
This can be due to the fact that directly translating a quantified formula into a quantifier-free  by introducing a vector of variables for each boolean possibility will blow up the size of the formula and may not be solvable by a SAT solver.

@@

In our synthesis method, we take a VTS $G =
(\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$ as input.
%
We allow activity functions not to be specified.
%
We will use the same encoding for the VTS as discussed in the QBF subsection.
%
We construct the following constraints to encode the available information
about $G$.
%
We encode both the present and the absent components in $G$.
%
Later, the constraints will help us encode the synthesis problems.
%
\begin{align*}
\texttt{PresentE} &= \land \{e_{i,j,q,m}|(i,M_1,j),...,(i,M_{q'},j) \in \edges \land q \leq q' \land m \in M_q \}\\
\texttt{PresentN} &= \land \{n_{i,m}| m \in \nlabel(i) \land i \in \nodes \}\\
\texttt{PresentP} &= \land \{p_{\{m_1,m_2,m_3,m_4\}}| \{m_1,m_2,m_3,m_4\} \in \pairs \} \\
\texttt{KnownActiveN} &= \land \{ a_{i,m} = \texttt{NodeFun}_m (n_{i,1},\dots,n_{i,\mu}) | \nodef_m \text{ is defined.} \} \\
\texttt{KnownActiveE} &= \land \{ b_{i,j,q,m} = \texttt{EdgeFun}_m(e_{i,j,q,1}, .., e_{i,j,q,\mu})
| \edgef_m \text{ is defined.} \} \\
\texttt{PresentCons} & = \texttt{PresentE} \land \texttt{PresentN} \land 
\texttt{PresentP} \land \texttt{KnownActiveE} \land \\
&  \texttt{KnownActiveN}
\end{align*}
We also collect the variables that are not set to true in $\texttt{PresentCons}$.
\begin{align*}
\texttt{AbsentELabel} &=
\{e_{i,j,q,m}|(i,M_1,j),...,(i,M_{q'},j) \in \edges \land 
0 < q \leq q' \land m \not\in M_{q} \}\\
\texttt{AbsentE} &= \{e_{i,j,q}|(i,M_1,j),...,(i,M_{q'},j) \in \edges \land 
q' < q \leq \pi \}\\
\texttt{AbsentNLabel} &= \{n_{i,m}| m \not\in \nlabel(i) \land i \in \nodes \}\\
\texttt{AbsentP} &= \{p_{\{m_1,m_2,m_3,m_4\}}| {\{m_1,m_2,m_3,m_4\}} \not\in \pairs \} \\
\texttt{UnknownActive} &=  \Land \{ a_{i,m} = \nodef_m (n_{i,1},\dots,n_{i,\mu}) | \nodef_m \text{ is undefined.} \} \union \\
&\quad\quad\quad  \{ b_{i,j,q,m} = \edgef_m(e_{i,j,q,1}, .., e_{i,j,q,\mu})
| \edgef_m \text{ is undefined.} \}
\end{align*}
We have defined $\texttt{AbsentELabel}$, $\texttt{AbsentE}$, $\texttt{AbsentN}$, and
$\texttt{AbsentP}$
as sets.
%
They will be converted into formulas depending
on the different usage in the synthesis problems.

>>>>

\subsubsection{Constraints for stability condition}
%
We use \texttt{NOracle} to represent the required path to encode the stability condition in VTS.
%
A naive implementation is shown in~\ref{eq:reachbmc1} where \texttt{NOracle}$(m, i, j, w)$ represent the presence of a $m$-$path$ from node $i$ to node $j$ of maximum $w$ length.
%there is a edge between $i^{\prime \prime}$ and01
%$(i \neq i^{\prime \prime})$ with $m$ present, and j is $m$-reachable from $i'$ in less than $p-1$ steps.
%
% \begin{align}
%   NOracle (m, i, j, w) = \exists \, path[h1, ..., hn]: \\ 2 \leq |path| \leq  w + 2  \, \land h1 = i \, \land  hn = j \, \land \\ 
%   \bigvee\limits_{k = 1}^{|path|- 1} \bigvee\limits_{q}  E[hk] [hk+1][q][m] 
%   \tag{S1}\label{eq:reach1}
% \end{align}
%The NOracle can be thought of as a disjunction of $m$-path of increment length. 
%
It performs enumeration of every $m$-$path$, $m$-${paths}_i^j$, of at most $w$ length. 
%
Starting with exploring $m$-${paths}_i^j$ of length 1, a direct edge between $i$ and $j$ with molecule $m$ present on it, until every $m$-${paths}_i^j$ of length $w-1$ is explored.
%by performing series of node hops till final node destination j is reached.
%Starting with $m$-${paths}_i^j$ of length 1 which is a direct edge between $i$ and $j$ with molecule $m$ present on the edge or performing series of node hops till final node destination j is reached.

@@

\subsection{Universal condition}
\label{enc:univ-cond}
%
Recall that the specification for universal condition~\ref{eq:univcond} requires quantifier alternation. 
%
Most SAT solvers can not efficiently handle the quantifier in the formulas. 
% 
Although, it is possible to encode the universal condition~\ref{eq:univcond} as a SAT or SMT problem.
%
There is always a danger of a blow-up in the formula size, making it harder for the SAT solver to handle. 
%Which may blow up the formula size for the solver to solve.
% 
As an experiment, we have implemented a direct SMT version of the universal problem extending the current SMT encodings.
%
The basic idea is to conjunct the formula for each combination value of the quantified variable. 
%

\input{satqbf-table}
%
The table~\ref{tab:satqbf-graph}, shows the search results of this encoding for the universal condition with 5 nodes and 3-connected graphs. 
%
The SMT extended version was not able to scale it to the desired results and either timed-out (T/O) or ran out of memory (M/O).
%
This can be due to the fact that directly translating a quantified formula into a quantifier-free  by introducing a vector of variables for each boolean possibility will blow up the size of the formula and may not be solvable by a SAT solver.

@@

In our synthesis method, we take a VTS $G =
(\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$ as input.
%
We allow activity functions not to be specified.
%
We will re-use the QBF encodings for the input VTS as described in the Section~\ref{enc:univ-cond}.
%
We construct the following constraints to encode the available information
about $G$.
%
We encode both the present and the absent components in $G$.
%
Later, the constraints will help us encode the synthesis problems.
%
\begin{align*}
\texttt{PresentE} &= \land \{e_{i,j,q,m}|(i,M_1,j),...,(i,M_{q'},j) \in \edges \land q \leq q' \land m \in M_q \}\\
\texttt{PresentN} &= \land \{n_{i,m}| m \in \nlabel(i) \land i \in \nodes \}\\
\texttt{PresentP} &= \land \{p_{\{m_1,m_2,m_3,m_4\}}| \{m_1,m_2,m_3,m_4\} \in \pairs \} \\
\texttt{KnownActiveN} &= \land \{ a_{i,m} = \texttt{NodeFun}_m (n_{i,1},\dots,n_{i,\mu}) | \nodef_m \text{ is defined.} \} \\
\texttt{KnownActiveE} &= \land \{ b_{i,j,q,m} = \texttt{EdgeFun}_m(e_{i,j,q,1}, .., e_{i,j,q,\mu})
| \edgef_m \text{ is defined.} \} \\
\texttt{PresentCons} & = \texttt{PresentE} \land \texttt{PresentN} \land 
\texttt{PresentP} \land \texttt{KnownActiveE} \land \\
&  \texttt{KnownActiveN}
\end{align*}
We also collect the variables that are not set to true in $\texttt{PresentCons}$.
\begin{align*}
\texttt{AbsentELabel} &=
\{e_{i,j,q,m}|(i,M_1,j),...,(i,M_{q'},j) \in \edges \land 
0 < q \leq q' \land m \not\in M_{q} \}\\
\texttt{AbsentE} &= \{e_{i,j,q}|(i,M_1,j),...,(i,M_{q'},j) \in \edges \land 
q' < q \leq \pi \}\\
\texttt{AbsentNLabel} &= \{n_{i,m}| m \not\in \nlabel(i) \land i \in \nodes \}\\
\texttt{AbsentP} &= \{p_{\{m_1,m_2,m_3,m_4\}}| {\{m_1,m_2,m_3,m_4\}} \not\in \pairs \} \\
\texttt{UnknownActive} &=  \Land \{ a_{i,m} = \nodef_m (n_{i,1},\dots,n_{i,\mu}) | \nodef_m \text{ is undefined.} \} \union \\
&\quad\quad\quad  \{ b_{i,j,q,m} = \edgef_m(e_{i,j,q,1}, .., e_{i,j,q,\mu})
| \edgef_m \text{ is undefined.} \}
\end{align*}
We have defined $\texttt{AbsentELabel}$, $\texttt{AbsentE}$, $\texttt{AbsentN}$, and
$\texttt{AbsentP}$
as sets.
%
They will be converted into formulas depending
on the different usage in the synthesis problems.

### 2019-03-10 papers/journal/model.tex
Improvement in Univ condition section 6.3

\ashu{Should be written bio people}
Let us discuss rough sizes of the system and expected behaviors.
%
A typical VTS may have about 10 compartments.
%
%Across eukaryotes, there are 20 broad SNARE varieties, though individual species can
%contain higher numbers (humans have 41)~\cite{kloepper2007elaborate}..
%
The VTS may transport a large number of molecules but the molecules that
are \todo{SA: commented line above} relevant of the control of VTS are about 50~\cite{kloepper2007elaborate}.
%
According to the definition the activity of molecules can be controlled by
the presence of all other molecules.
%
However, in practice the activity of a molecule is controlled by
the presence of ~2-3 other molecules.

>>>>

\ashu{Should be written bio people}
Let us discuss rough sizes of the system and expected behaviors.
%
A typical VTS may have about 10 compartments.
%
%Across eukaryotes, there are 20 broad SNARE varieties, though individual species can
%contain higher numbers (humans have 41)~\cite{kloepper2007elaborate}..
%
The VTS may transport a large number of molecules but the molecules that
are relevant of the control of VTS are about 50~\cite{kloepper2007elaborate}.
%
According to the definition the activity of molecules can be controlled by
the presence of all other molecules.
%
However, in practice the activity of a molecule is controlled by
the presence of ~2-3 other molecules.
